Necessity is the mother of genetic invention

### 2017-11-03 ms.tex
finished with MWF comments of Nov 2

\begin{equation}
\begin{aligned}
\frac{\partial \bar{\omega}^*}{\partial C_k} = 
v_k D_k (u\tr[M]_k - u_k) = \\
D_k (\sum_i{u_i M_{i,k} v_k} - u_k v_k).
\end{aligned}
\end{equation}
Here, $u_k$ can be regarded as the \emph{reproductive value} of type $k$ (see~\nameref{sec:AppC}), which gives the relative contribution of type $k$ to the long-term population.
Subsequently, $u_k v_k$ is the \emph{ancestor frequency} \citep{Hermisson2002} of type $k$, which is the fraction of the equilibrium population descended from type $k$.
The sum $\sum_i{u_i M_{i,k} v_k}$ can be similarly interpreted as the fraction of the equilibrium population descended from individuals that transitioned from type $k$ to another type (via the $k$ column of the transition matrix $M$).

@@

\section*{Appendices}
% Appendix A
\subsection*{Appendix A}\label{sec:AppA}

@@

% Appendix B
\subsection*{Appendix B}\label{sec:AppB}

\citet{Caswell1978} presented a \emph{formula for the sensitivity of the population growth rate to changes in life history parameters}.
In this formula, the \emph{population growth rate} is the leading eigenvalue of the population transformation matrix $T$, the \emph{life history parameters} are elements of $T$, and the \emph{sensitivity} is the derivative of the former with respect to the latter.
This is a useful formula~\citep{Hermisson2002,Ram2012}, and therefore we reproduce it here.

\begin{caswell}
Let $T$ be a matrix with eigenvalue $\lambda$ and left and right eigenvectors $u$ and $v$ such that $\sum{v_k}=1$ and $u\tr v = \sum{u_k v_k} = 1$.
Then the sensitivity of $\lambda$ to changes in any element $t$ of the matrix $T$ is
\begin{equation}\label{eq:Caswells_formula}
\frac{\partial \lambda}{\partial t} = 
u\tr \frac{\partial T}{\partial t} v
\end{equation}
\end{caswell}

@@

% Appendix C
\subsection*{Appendix C}\label{sec:AppC}
\begin{frv}
Let $M$ be a stochastic matrix and $D$ be a diagonal matrix with positive diagonal elements. 
The elements of the left \emph{Perron} eigenvector $u$ of the matrix $MD$ can be regarded as \emph{Fisher's reproductive values}~\citep{Fisher1930,Grafen2006}.
\end{frv}

\emph{Fisher's reproductive values} can be understood as follows.
Let us look at the dynamics not of frequencies but of absolute population sizes such that the vector of the number of individuals of each type at time $t$ is $n(t)$ and $f_k(t) = n_k(t) / \sum_i{n_i(t)}$.
The dynamics are
\begin{equation}
n(t) = (MD)^t n(0).
\end{equation}

@@

To further evaluate, we write $MD$ in its Jordan canonical form 
\begin{equation}
A = V \Lambda U\tr,
\end{equation}
where $V$ is the matrix of right column eigenvectors of $MD$,
$U\tr$ is the matrix of left row eigenvectors of $MD$,
$V U\tr = U\tr V = I$, 
and $\Lambda$ is the diagonal matrix of eigenvalues of $A$ 
(for a non-generic set of matrices $M$, the geometric and algebraic multiplicities of the eigenvalues of $MD$ differ, and $\Lambda$ will not be a diagonal matrix, a case we can ignore).

@@

Now take the long term limit $t \goesto \infty$.
By assumption, $MD$ is irreducible, so $\lambda_i < \lambda_1$ for all $i > 1$.
Therefore,
\begin{equation}
\lim_{t \goesto \infty} \Big( \frac{\lambda_k}{\lambda_1} \Big)^t = 0
\end{equation}
for all $k > 1$, and
\begin{equation}
\lim_{t \goesto \infty}\frac{N(j,t)}{N(k,t)} =
\frac{e\tr V (e_1 e_1\tr) U\tr e_j} { e\tr V (e_1 e_1\tr) U\tr e_k} = 
\frac{ e\tr v u_j} { e\tr v  u_k} =
\frac{u_j} {u_k}.
\end{equation}

The vector $u$ is the left \emph{Perron} eigenvector of $MD$, and $u_k$ is element $k$ of $u$.
This is why the value $u_k$ can be interpreted as the \emph{reproductive value} of type $i$: it is a weight for the size of the population generated by a single individual of type $k$.

If we begin with a population at the equilibrium distribution $v$,
and ask what fraction of long-term descendants descended from type $k$ at that time,
we weight the equilibrium frequency $v_k$ by the reproductive value $u_k$, to get $u_k v_k$.
$\{u_k v_k\}_{Ik}$ is a probability distribution, since
\begin{equation}
\sum_k {u_k v_k} = u\tr v = 1.
\end{equation}
\citet{Hermisson2002} called this distribution the \emph{ancestor} or \emph{ancestral distribution}.

>>>>

\begin{equation}
\begin{aligned}
\frac{\partial \bar{\omega}^*}{\partial C_k} = 
v_k D_k (u\tr[M]_k - u_k) = \\
D_k (\sum_i{u_i M_{i,k} v_k} - u_k v_k).
\end{aligned}
\end{equation}
Here, $u_k$ can be regarded as the \emph{reproductive value} of type $k$ (see~\nameref{sec:AppC}), which gives the relative contribution of type $k$ to the long-term population.
Consequently, $u_k v_k$ is the \emph{ancestor frequency} \citep{Hermisson2002} of type $k$, which is the fraction of the equilibrium population descended from type $k$.
The sum $\sum_i{u_i M_{i,k} v_k}$ can be similarly interpreted as the fraction of the equilibrium population descended from individuals that transitioned from type $k$ to another type (via the $k$ column of the transition matrix $M$).

@@

\section*{Appendices}
% Appendix A: mean fitness
\subsection*{Appendix A}\label{sec:AppA}
\textbf{TODO} reference to Fisher 1930 or Fisher 1958 (Caswell 1978 cites the latter in pg. 217 just above eq. 4).

@@

% Appendix B: Caswell
\subsection*{Appendix B}\label{sec:AppB}

\citet{Caswell1978} gave a \emph{formula for the sensitivity of the population growth rate to changes in life history parameters}.
In this formula, the \emph{population growth rate} is the leading eigenvalue of the population transformation matrix $T$, the \emph{life history parameters} are elements of $T$, and the \emph{sensitivity} is the derivative of the former with respect to the latter.
This is a useful formula~\citep{Caswell1978,Hermisson2002,Ram2012}, and therefore we reproduce it here.

\begin{caswell}
Let $T$ be a non-negative matrix with leading eigenvalue $\lambda$ and left and right eigenvectors $u$ and $v$ such that $\sum{v_k}=1$ and $u\tr v = \sum{u_k v_k} = 1$.
Then the sensitivity of $\lambda$ to changes in any element $t$ of the matrix $T$ is
\begin{equation}\label{eq:Caswells_formula}
\frac{\partial \lambda}{\partial t} = 
u\tr \frac{\partial T}{\partial t} v
\end{equation}
\end{caswell}

@@

% Appendix C: Fisher's reproductive value
\subsection*{Appendix C}\label{sec:AppC}
\begin{frv}
Let $M$ be a stochastic matrix and $D$ be a diagonal matrix with positive diagonal elements. 
The elements of the left \emph{Perron} eigenvector $u$ of the matrix $MD$ can be regarded as \emph{Fisher's reproductive values}~\citep{Fisher1930,Grafen2006}.
\end{frv}

\emph{Fisher's reproductive values} can be understood as follows.
Consider the dynamics not of frequencies but of absolute population sizes such that the vector of the number of individuals of each type at time $t$ is $n(t)$ and $f_k(t) = n_k(t) / \sum_i{n_i(t)}$.
The dynamics are
\begin{equation}
n(t) = (MD)^t n(0).
\end{equation}

@@

Now write $MD$ in its Jordan canonical form 
\begin{equation}
A = V \Lambda U\tr,
\end{equation}
where $V$ is the matrix of right column eigenvectors of $MD$,
$U\tr$ is the matrix of left row eigenvectors of $MD$,
$V U\tr = U\tr V = I$, 
and $\Lambda$ is the diagonal matrix of eigenvalues of $A$ 
(for a non-generic set of matrices $M$, the geometric and algebraic multiplicities of the eigenvalues of $MD$ differ, and $\Lambda$ will not be a diagonal matrix, a case we can ignore).

@@

Now take the long termlimit $t \goesto \infty$.
By assumption, $MD$ is irreducible, so $\lambda_i < \lambda_1$ for all $i > 1$.
Therefore,
\begin{equation}
\lim_{t \goesto \infty} \Big( \frac{\lambda_k}{\lambda_1} \Big)^t = 0
\end{equation}
for all $k > 1$, and
\begin{equation}
\lim_{t \goesto \infty}\frac{N(j,t)}{N(k,t)} =
\frac{e\tr V (e_1 e_1\tr) U\tr e_j} { e\tr V (e_1 e_1\tr) U\tr e_k} = 
\frac{ e\tr v u_j} { e\tr v  u_k} =
\frac{u_j} {u_k}.
\end{equation}

The vector $u$ is the left \emph{Perron} eigenvector of $MD$, and $u_k$ is element $k$ of $u$.
This is why the value $u_k$ can be interpreted as the \emph{reproductive value} of type $i$: it is a weighting for the size of the population generated by a single individual of type $k$.

If we begin with a population at the equilibrium distribution $v$,
and ask what fraction of long-term descendants descended from type $k$ at that time,
we weight the equilibrium frequency $v_k$ by the reproductive value $u_k$, to get $u_k v_k$.
$\{u_k v_k\}_{k}$ is a probability distribution, since
\begin{equation}
\sum_k {u_k v_k} = u\tr v = 1.
\end{equation}
\citet{Hermisson2002} called this distribution the \emph{ancestor} or \emph{ancestral distribution}.
